Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-20 cav2017/cav2017.tex
first round of cleanup
Removed observer and floating point from main text

We consider physical models expressed as ordinary differential equations (ODEs) with full state information:  
%
\begin{align}
\label{eq:ode}
\dot{x}(t) = Ax(t)+ B u(t)\ \ :\ \ x \in \mathbb{R}^{n}, u \in \mathbb{R}^m, A \in \mathbb{R}^{n \times n}, B \in \mathbb{R}^{n \times m}
\end{align}
where $A$ and $B$ are matrices that fully specify the continuous plant with the initial states $x(0)$. 
Eq.~\eqref{eq:ode} is soundly discretised (as shown in the Appendix) into
%
\begin{align}
\label{eq:plant}
x_{k+1} = A_d x_k+ B_d u_k 
y_{k} = C_d x_k+ D_d u_k,
\end{align}
with initial state $x_{0}=x(0)$.

@@

@@

Let $\mathbb{R} \langle I,F \rangle$ be a fixed point domain with $I$ bits representing the integer part and $F$ bits
representing the decimal part. We will extend this notation to $\mathbb{R} \langle I,F,E \rangle$ to add a number of
leading or trailing bits when the domain requires prescaling (ie the maximum is smaller than $0.5$ or the minimum
is greater than $2$).
Let also $\mathbb{R}_* \langle M,E \rangle$ be a floating point domain with $M$ bits
mantissa and $E$ bits exponent. This domain corresponds to 
$$\mathbb{R}_* \langle M,E \rangle = \bigcup_i \mathbb{R} \langle 1,M-1,E_i \rangle : 2^{1-E} \leq E_i  \leq 2^{E-1}$$
This means that the distance between any two numbers in the domain depends on the
location of those numbers.
Since we've shown that a floating point domain is a union of a number of fixed point domains, we will continue our
reasoning on fixed point only. Our reasoning will apply to the floating point domain by extension.
The following approximation errors will arise:
\begin{enumerate}
\item {\bf Truncation:}
Let $\mathcal{F}\langle I,F \rangle(\cdot) : \mathbb{R} \rightarrow \mathbb{R}\langle I,F \rangle, \tilde x=\mathcal{F}\langle I,F \rangle(x) = x-\delta : \delta=x\ \%_F\  c\tilde x$, where $\%_F$ is the modulus operation performed on the last bit of the representation ($2^{-F}$).
The number $\delta$ corresponds to the truncation error of the representation and it will propagate across operations.
\item {\bf Rounding:}
%% Every time an operation is performed on a number in the $\mathbb{R} \langle I,F,E \rangle$ domain, an error may be
%% introduced due to truncation of the result.
Let us define $c\langle i,f,e \rangle \in \mathbb{R}\langle I,F,E\rangle$, where $i \in [-2^{I-1}\ 2^{I-1}],f \in [0 1),e \in [-2^{E-1}\ 2^{E-1}]$ are its assigned values. The smallest number that can be represented for any given exponent is $c_{min(e)}=c\langle 0,2^{-F},e\rangle$.
The following errors appear in simple operations:
\begin{enumerate}
\item Addition/Substraction: $c\langle \cdot,e_1\rangle \pm c\langle \cdot,e_2\rangle=c\langle \cdot,max(e_1,e_2)\rangle + \delta_3 : |\delta_3| \leq (e_1==e_2) ? 0 : c_{min(max(e_2,e_1))}$ 
\item Multiplication: $c\langle \cdot,e_1\rangle * c\langle \cdot,e_2\rangle=c\langle \cdot,e_1+e_2\rangle + \delta_3 : |\delta_3| \leq c_{min(e_2+e_1)}$
\end{enumerate}
The bounds obtained for division are left to the reader as it will not be applied in this paper.

At this point it is worth mentioning that not all systems truncate equally. In the definition above $\delta$ may be positive
or negative, and even this decision can be made dependent on the sign of $x$. Common cases in computer systems are:
round downwards (\emph{ie} to $-\infty$), round upwards (\emph{ie} to $+\infty$), and round to zero.
Rounding errors are cumulative which means the overall error will statistically increase with every operation, thus algorithms
performing fewer operations can be more precise in this respect than iterative ones. 
\item {\bf Overflow:}
  %Another source of error introduced by Finite Word Length representations is that of overflow.
  The finite word length set a
maximum representable number in the domain ($\pm 2^{I-1}+1-2^{-F}$), which means that any number outside this range
cannot be accurately represented by the domain (\emph{eg} $\delta=x-2^{I-1}+1-2^{-F}$). In the case of floating point, a special
case for infinity has been introduced, which addresses some of this issues, but once the error has gone to infinity, the numbers
become meaningless. It is the responsibility of software designers and tools to account for overflows and either supply saturation
functions to prevent them or raise alarms in their presence. In the case of verification, the latter is seen as a safety problem.
\item {\bf Null/Zeno behaviour:}
Whilst this work does not explore the effect of Zeno behaviours, it is worth noting that this effect may also be caused by rounding
errors in floating point programs. The reason for this is that if an operation contains two numbers which belong to disjoint sub-domains
(\emph{ie} neither number can be represented in the domain of its counterpart), the resulting operation may become a no-op/skip,
subsequently destabilizing the program.
\end{enumerate}

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Fixed word length effects on matrix multiplications}
\label{sec:cof_fwl}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Thus far we have presented a verification procedure for a digitalized plant in a continuous space (\emph{ie} the
range of the quatizers is still $\mathbb{R}$. The next step we must reaize is that the controller and the observer
exist in a digital program, hence their domain is $\mathbb{R}\langle I,F,E\rangle$. This has two separate effects
on the model. If we recall equation \eqref{eq:to_of}, the matrices of the observer and the controller  are not necessarily those of the plant.
In fact we have $\mat{A},\mat{B},\mat{C} \in \mathbb{R}^*$, $\mat{A}_o,\mat{B}_o,\mat{C}_o, \mat{K}_d, \mat{L}_d \in \mathbb{R}\langle I,F,E\rangle^*$
Where $*$ represents the appropriate dimension for each matrix.
Unfortunately this means that although combined matrices such as $\mat{A}_t$ are in the reals, they suffer from
the fixed word restrictions in their composing matrices.
The second effect corresponds to the rounding errors in the calculation of $\hat{x}_{k+1}, u_k$. These will introduce a new source of quantiztion noise which we may denote $\vec{\nu}_3 \in [-q_3\ \ q_3]$.
The estimation of $\hat{\vec{x}}$ corresponds to the sum of three matrix-vector multiplications (we assume $\mat{C}_o$ in Objective form). Each of these has a coefficient-wise error bound $c_{min(max(e_{j1}+e_{j2}))}$. Hence the overall $\overline{\delta}_{\hat{x}}=3nc_{min(max(e_{j1}+e_{j2}))}$. The error in $\vec{u}$ corresponds of the multiplication of $\mat{K}_d\hat{\vec{x}}$, hence $q_3=\overline{\delta}_u=nc_{min(max(e_{k1}+e_{k2}))}+\sum_{k=1}^nc_k\overline{\delta}_{\hat{x}})=3n|\mat{K}|_1c_{min(max(e_{j1}+e_{j2}))}+nc_{min(max(e_{k1}+e_{k2}))}$

@@

%\todo{The safety specification is defined over the state space, not the output. Further, please distinguish the spec from the input constraints. }

@@

Let us recall the formulas used for single-input and single-output (SISO) systems 
in sections \ref{sec:reachable} and \ref{sec:observable}.
A controllable and observable system will have the following dynamics:
\textcolor{red}{Dario: Do we really need to describe all these equations here?
Should we simply cite them from a given reference?}
%
\begin{align}
\label{eq:observer_LTI_cf}
\left [\begin{array}{c}\vec{x}\\ \hat{\vec{x}}_e \end{array}\right]_{k+1}
=&\left [\begin{array}{cc}\mat{A}_{tp}&\mat{B}_{to}\\ \mat{0}&\mat{A}_{to}\end{array}\right]
\left [\begin{array}{c}\vec{x}\\ \hat{\vec{x}}_e \end{array}\right]_k
+\mat{B}_{tp} \vec{r}_k
+\mat{B}_{tn}\left [\begin{array}{c}\vec{\nu}_2\\ \vec{\nu}_1\end{array}\right]\\
\vec{y}_k=&\mat{C}_{t}\vec{x}_k + \mat{D}_{d}\vec{u}_k\nonumber
\end{align}
\begin{align*}
\mat{A}_{tp}=&\mat{T}\mat{A}_d\mat{T}^{-1}-\mat{T}\mat{B}_d\mat{K}_d\mat{T}^{-1}&
\mat{A}_{to}=&\mat{T}\mat{A}_d\mat{T}^{-1}-\mat{L}_d\mat{C}_d\mat{T}&
\mat{A}_t=&\left [\begin{array}{cc}\mat{A}_{tp}&\mat{B}_{to}\\ \mat{0}&\mat{A}_{to}\end{array}\right]\\
\mat{B}_{to}=&\mat{T}\mat{B}_d\mat{K}_d\mat{T}^{-1}&
\mat{B}_{tp}=&\left[\begin{array}{c}\mat{T}\mat{B}_d\mat{k}_r\\\mat{0}\end{array}\right]&
\mat{B}_{tn}=&\left[\begin{array}{cc}\mat{T}\mat{B}_d\mat{k}_r&\mat{0}\\\mat{0}&-\mat{L}_d\end{array}\right]\\
\mat{T}=&\mat{R}_{cf}\mat{R}^{-1}&
\mat{L}=&\mat{W}_o^{-1}\mat{W}_{of} \left[ \begin{array}{c}p_1-a_1\\p_2-a_2\\p_n-a_n\end{array}\right]&
\mat{C}_t=&\mat{C}_d\mat{T}
\end{align*}

@@

Careful examination of these matrices reveals that because the $\mat{A}_t$ is 
upper block diagonal, the spectrum of $\mat{A}_t$ is equal to the union of the 
spectrum of $\mat{A}_{tp}$ and $\mat{A}_{to}$, which can be calculated separately. 
Additionally, these correspond to the roots of the following polynomials:
%
\begin{align*}
\left(z^n+\sum_{i=1}^p{(a_i-k_i)z^{n-i}}\right) \left(z^n+\sum_{i=1}^p{(p_i-a_i)z^{n-i}}\right)
\end{align*}

@@

Notice that in both our polynomials $c_0=1$ and we only use reals.
Now that we have established convergence, we may examine the safety 
specification of the state space. In the case of SISO systems, 
this is given by a range of the input and a range of the output, 
which are considered safe ($\vec{u}_k \in [\underline{u}\ \ \overline{u}] \wedge \vec{y}_k \in [\underline{y}\ \ \overline{y}]$). 
Whilst there are other properties, we may include in the verification, 
we will solely focus on these for simplicity of the explanation.
We will assume that we do not want saturation in our system, 
hence the control loop has only one mode. We have also identified 
the quantization noise as a bounded non-deterministic input.
First we must translate the safety specification into the state-space. 
Replacing on the second part of equation \eqref{eq:observer_LTI_cf} we obtain
%
\begin{equation}
\Psi := \underline{y}-\overline{\mat{B}\vec{u}}<\mat{C}_t\vec{x}_k<\overline{y}-\underline{\mat{B}\vec{u}}
\label{eq:of_bounds}
\end{equation}
Note that we use the bounds of the input after the transformation by $\mat{B}$ 
since the matrix may reverse the bounds. Eq.~\eqref{eq:of_bounds} can be solved 
using linear programming as it describes a polyhedral region in $\mathbb{R}^n$.
The second step is evaluating the progression of the system over time to calculate 
$\vec{x}_k$ to validate against this specification. For this step we require 
an initial set from which the system progresses. Obviously, this set $\vec{x}_0$ 
must satisfy the specification; otherwise, the system will be unsafe to begin with.
We accept a specification in the form $\mat{E}\mat{T}\vec{x}_0<\mat{f}$. 
The presence of $\mat{T}$ is because this will typically be given in the original 
state-space. Accelerating \eqref{eq:observer_LTI_cf}, we obtain:
%
\begin{align}
\label{eq:acc_observer_LTI_cf}
\left [\begin{array}{c}\vec{x}\\ \hat{\vec{x}}_e \end{array}\right]_k
=&\mat{A}_t^k
\left [\begin{array}{c}\vec{x}\\ \hat{\vec{x}}_e \end{array}\right]_0
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tp} \vec{r}_i
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tn}\left [\begin{array}{c}\vec{\nu}_2\\ \vec{\nu}_1\end{array}\right] :
 \left [\begin{array}{cc}\mat{E}\mat{T}&\mat{0}\end{array}\right] \left [\begin{array}{c}\vec{x}\\ \hat{\vec{x}}_e \end{array}\right]_0<\mat{f}
\end{align}
which would still require us to verify the system for every $k$ up to infinity. 
Instead we use abstract acceleration~\cite{cattaruzza2015unbounded}, which translates the formula into
%
\begin{align}
\label{eq:aa_observer_LTI_cf}
\hat{X}^\#
=&\mathcal{A}_t X_0+\mathcal{B}_{tp} R + \mathcal{B}_{tn} N\\
X_0 =&\left \{ \left[\begin{array}{c}\vec{x}\\ \hat{\vec{x}}_e \end{array}\right] :
 \left [\begin{array}{cc}\mat{E}\mat{T}&\mat{0}\end{array}\right] \left [\begin{array}{c}\vec{x}\\ \hat{\vec{x}}_e \end{array}\right]_0<\mat{f} \right\}\nonumber\\
 R =&\{r : \mat{E}_r < \vec{f}_r \}\nonumber\\
 N=&\left \{ \left[\begin{array}{c}\vec{\nu}_2\\ \vec{\nu}_1\end{array}\right] : \nu_1 \in \left[-\frac{q1}{2}\ \ \frac{q1}{2}\right] \wedge \nu_2 \in \left[-\frac{q2}{2}\ \ \frac{q2}{2}\right]  \right \}\nonumber
\end{align}
%
where $\mathcal{A}_t=\bigcup_{k=1}^\infty \mat{A}^k, \mathcal{B}_{tp}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tp}, \mathcal{B}_{tn}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tn}$ are abstract matrices for the system in \eqref{eq:observer_LTI_cf}. It is worth noting that $r$ is a non-deterministic parametric input (\emph{ie} $\forall k>0, r_{k+1} =r_k$) whereas the set N is composed of non-deterministic variable inputs.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Controller synthesis}\label{sec:cof_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Now that we have explained the verification process, we may proceed to describe the synthesis procedure, which is illustrated in figure \ref{fig:CEGARIS}.
\begin{enumerate}
\item We start by preprocessing the dynamics. This can be computationally intensive and we don't want it to be part of the CEGIS loop.
\begin{enumerate}
\item Calculate $\mat{A}_o,\mat{B}_o,\mat{C}_o$ by rounding down the coefficients $*_{oij}=\mathcal{F}\langle I,F,E\rangle (*_{ij}) : *=\{a,b,c\}$
\item Calculate $\mat{T}$ and $\mat{W}_o^{-1}\mat{W}_{of}$ using equations \eqref{eq:rncf}\eqref{eq:rcf}\eqref{eq:to_cf}\eqref{eq:wnof}\eqref{eq:wof}
\item Extract the Plant's characteristic polynomial coefficients from the last row of $P_a=(\mat{T}\mat{A}\mat{T}^-1)_{n*}$.
\item Calculate the noise set $N$ from the quantizer resolutions and estimated round-off errors ($\frac{q_1}{2},\frac{q_2}{2},q_3$)
\item Produce an initial abstraction which corresponds to an unconstrained solution space.
\end{enumerate}
\item Next we synthesise a candidate controller and observer. Let $\mathcal{J}\langle I,F,E \rangle (P,\mat{T})$ be a program describing Jury's method.
\begin{enumerate}
\item Select a controller $\tilde{\mat{K}} \in \mathbb{R}\langle I,F,E\rangle^n$. This is achieved by solving a SAT formula $\mathcal{J}_K=\mathcal{J}\langle I,F,E \rangle (P_{a-k},\mat{T})$ that satisfies Jury's criteria for $P_{a-k}=z^n+\sum_{i=1}^n (a_i-k_i) z^{n-i} : k_i \in \mat{K} \wedge \mat{K}=\tilde{\mat{K}}\mat{T}$.
\item Select the observer error characteristic polynomial $\tilde{\mat{L}} \in \mathbb{R}\langle I,F,E\rangle^n$. 
This is achieved by solving a SAT formula $\mathcal{J}_L=\mathcal{J}\langle I,F,E \rangle (P_p,\mat{W}_o^{-1}\mat{W}_{of})$ 
that satisfies Jury's criteria for 
$P_p=z^n+\sum_{i=1}^n p_i z^{n-i} : p_i \in \vec{P}= (\mat{W}_o^{-1}\mat{W}_{of})^-1 \tilde{\mat{L}} 
+ [\begin{array}{cccc}a_1&a_2&\hdots&a_n\end{array}]^T$.
\item If there is no candidate solution we return UNSAT and exit the loop.
\end{enumerate}
\item Once we have a candidate solution, we perform a safety verification using abstract acceleration.
\begin{enumerate}
\item Calculate the abstract matrices $\mathcal{A}_t, \mathcal{B}_{tp},\mathcal{B}_{tn}$ for the candidate closed loop solution.
\item Evaluate $\hat{X}^\# \models \Psi$. If the verification holds we have a solution, and exit the loop.
\end{enumerate} 
\item If the candidate solution is not valid we refine the abstraction used by the synthesiser
\begin{enumerate}
\item Find the constraints that invalidate the property. These are the rows of $\mat{E} \in \Psi$ that intersect/segment $\hat{X}^\#$. Transforming these rows into the eigenspace of $\mat{A}_t$, and applying backward acceleration to obtain a set of counterexamples for the eigenvalues, which we define as $\Psi(\Lambda)$. This is a constraint in the spectrum (\emph{ie} frequency response transfer function) of the closed loop dynamics. 
\item Transform the soft constraints on the spectrum into soft constraints on the coefficients of the characteristic polynomial. 
$z^n+\sum_{i=1}^n[a_i-k_i]z^{n-i}=\prod_{i=1}^n (z-[\lambda_i]) : |[\lambda_i]|<1 \wedge [\lambda_i] \models \Psi_{\Lambda}$.
\item Pass the refined abstraction with the new contraints to the synthesiser, as well as a counterexample plant coefficients ($P_a$) and repeat the loop again. We note that the Abstraction refinement is a soft constraint on the proposed model which can increase or decrease the parameter space at any time, whereas the concrete counterexample passed to the synthesiser is monotonically decreasing. The first one improves speed of the tool whereas the second ensures convergence.

>>>>

We consider physical models expressed as ordinary differential equations (ODEs) with full state information:  
%
\begin{align}
\label{eq:ode}
\dot{x}(t) = Ax(t)+ B u(t)\ \ :\ \ x \in \mathbb{R}^{n}, u \in \mathbb{R}^m, A \in \mathbb{R}^{n \times n}, B \in \mathbb{R}^{n \times m}
\end{align}
where $A$ and $B$ are matrices that fully specify the continuous plant with the initial states $x(0)$. 
Eq.~\eqref{eq:ode} is soundly discretised (as shown in the Appendix) into
%
\begin{align}
\label{eq:plant}
x_{k+1} = A_d x_k+ B_d u_k. 
\end{align}
with initial state $x_{0}=x(0)$.

@@

\todo{HERE add a figure similar to Fig. \label{fig:observersystem} to explain digital features of controller.}

\todo{[The remainder of the section needs to be updated. ]}
The output and measurable of the plant is being sampled by an ADC, and since the full state of the
plant cannot be witnessed directly, the controller software includes an observer that 
estimates the state of the plant (see appendix for details on how such an observer
is constructed). In this work, we are assuming that all the states measurements are available for the state feedback purpose, and the state observer design may be disregarded.  
\todo{[Unclear what the remainder is for: ]}

@@

Let $\mathbb{R} \langle I,F \rangle$ be a fixed point domain with $I$ bits representing the integer part and $F$ bits
representing the decimal part. 
The following approximation errors will arise:
\begin{enumerate}
\item {\bf Truncation:}
Let $\mathcal{F}\langle I,F \rangle(\cdot) : \mathbb{R} \rightarrow \mathbb{R}\langle I,F \rangle, \tilde x=\mathcal{F}\langle I,F \rangle(x) = x-\delta : \delta=x\ \%_F\  c\tilde x$, where $\%_F$ is the modulus operation performed on the last bit of the representation ($2^{-F}$).
The number $\delta$ corresponds to the truncation error of the representation and it will propagate across operations.
\item {\bf Rounding:}
%% Every time an operation is performed on a number in the $\mathbb{R} \langle I,F,E \rangle$ domain, an error may be
%% introduced due to truncation of the result.
Let us define $c\langle i,f \rangle \in \mathbb{R}\langle I,F\rangle$, where $i \in [-2^{I-1}\ 2^{I-1}],f \in [0 1)$ are its assigned values. The smallest number that can be represented for any given exponent is $c_{min}=2^{-F}$.
The following errors appear in simple operations:
\begin{enumerate}
\item Addition/Substraction: in the case of fixed point arithmetic of numbers with the same domain, there are no rounding errors, otherwise the maximum error is given by the domain of the result $c\langle \cdot, F_1\rangle \pm c\langle \cdot , F_2\rangle= c\langle \cdot , F_3\rangle + \delta_3 :|\delta_3| \leq 2^{-F_3} \wedge F_3 \leq F_1,F_2$ 
\item Multiplication: $c\langle \cdot, F_1\rangle * c\langle \cdot,F_2\rangle=c\langle \cdot,F_3\rangle + \delta_3 : |\delta_3| \leq 2^{-F_3} \wedge F_3 \leq F_1,F_2$ 
\end{enumerate}
The bounds obtained for division are left to the reader as it will not be applied in this paper.

At this point it is worth mentioning that not all systems truncate equally. In the definition above $\delta$ may be positive
or negative, and even this decision can be made dependent on the sign of $x$. Common cases in computer systems are:
round downwards (\emph{ie} to $-\infty$), round upwards (\emph{ie} to $+\infty$), and round to zero.
Rounding errors are cumulative which means the overall error will statistically increase with every operation, thus algorithms
performing fewer operations can be more precise in this respect than iterative ones. 
\item {\bf Overflow:}
  %Another source of error introduced by Finite Word Length representations is that of overflow.
  The finite word length set a
maximum representable number in the domain ($\pm 2^{I-1}+1-2^{-F}$), which means that any number outside this range
cannot be accurately represented by the domain (\emph{eg} $\delta=x-2^{I-1}+1-2^{-F}$). In the case of floating point, a special
case for infinity has been introduced, which addresses some of this issues, but once the error has gone to infinity, the numbers
become meaningless. It is the responsibility of software designers and tools to account for overflows and either supply saturation
functions to prevent them or raise alarms in their presence. In the case of verification, the latter is seen as a safety problem.
\end{enumerate}

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Fixed word length effects on matrix multiplications}
\label{sec:cof_fwl}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Thus far we have presented a verification procedure for a digitalized plant in a continuous space (\emph{ie} the
range of the quatizers is still $\mathbb{R}$. The next step we must reaize is that the controller and the observer
exist in a digital program, hence their domain is $\mathbb{R}\langle I,F\rangle$. This has two separate effects
on the model. If we recall equation \eqref{eq:to_of}, the matrices of the observer and the controller  are not necessarily those of the plant.
In fact we have $\mat{A},\mat{B} \in \mathbb{R}^*$, $\mat{A}_o,\mat{B}_o,\mat{K}_d, \mat{L}_d \in \mathbb{R}\langle I,F\rangle^*$
Where $*$ represents the appropriate dimension for each matrix.
Unfortunately this means that although combined matrices such as $\mat{A}_t$ are in the reals, they suffer from
the fixed word restrictions in their composing matrices.
The second effect corresponds to the rounding errors in the calculation of $\hat{x}_{k+1}, u_k$. These will introduce a new source of quantiztion noise which we may denote $\vec{\nu}_3 \in [-q_3\ \ q_3]$.
The estimation of $\hat{\vec{x}}$ corresponds to the sum of three matrix-vector multiplications. Each of these has a coefficient-wise error bound $c_{min}$. Hence the overall $\overline{\delta}_{\hat{x}}=3nc_{min}$. The error in $\vec{u}$ corresponds of the multiplication of $\mat{K}_d\hat{\vec{x}}$, hence $q_3=\overline{\delta}_u=nc_{min}+\sum_{k=1}^nc_k\overline{\delta}_{\hat{x}}=(3|\mat{K}|_1+1)n2^{-F}$

@@

@@

Let us recall the formulas used for single-input and single-output (SISO) systems 
in sections \ref{sec:reachable} and \ref{sec:observable}.
A controllable and observable system will have the following dynamics:

%
\begin{align}
\label{eq:observer_LTI_cf}
\vec{x}_{k+1}=&\mat{A}_{t}\vec{x}_k+\mat{B}_{tp} \vec{r}_k+\mat{B}_{tn}\left [\begin{array}{c}\vec{\nu}_2\\ \vec{\nu}_1\end{array}\right]
\end{align}
\begin{align*}
\mat{A}_{t}=&\mat{T}\mat{A}_d\mat{T}^{-1}&
\mat{B}_{tp}=&\mat{T}\mat{B}_d&
\mat{B}_{tn}=&\mat{T}\mat{B}_d&
\mat{T}=&\mat{R}_{cf}\mat{R}^{-1}&
\end{align*}

@@

Careful examination of these matrices reveals that because the $\mat{A}_t$ is 
upper block diagonal, the spectrum of $\mat{A}_t$ is equal to the union of the 
spectrum of $\mat{A}_{tp}$ and $\mat{A}_{to}$, which can be calculated separately. 
Additionally, these correspond to the roots of the following polynomials:
%
\begin{align*}
\left(z^n+\sum_{i=1}^p{(a_i-k_i)z^{n-i}}\right)
\end{align*}

@@

Notice that in both our polynomials $c_0=1$ and we only use reals.
Now that we have established convergence, we may examine the safety 
specification of the state space. In the case of SISO systems, 
this is given by a range of the input and a range of the output, 
which are considered safe ($\vec{u}_k \in [\underline{u}\ \ \overline{u}] \wedge \vec{y}_k \in [\underline{y}\ \ \overline{y}]$). 
Whilst there are other properties, we may include in the verification, 
we will solely focus on these for simplicity of the explanation.
We will assume that we do not want saturation in our system, 
hence the control loop has only one mode. We have also identified 
the quantization noise as a bounded non-deterministic input.
First we must translate the safety specification into the state-space. 
Replacing on the second part of equation \eqref{eq:observer_LTI_cf} we obtain
%
\begin{equation}
\Psi := \underline{y}-\overline{\mat{B}\vec{u}}<\mat{C}_t\vec{x}_k<\overline{y}-\underline{\mat{B}\vec{u}}
\label{eq:of_bounds}
\end{equation}
Note that we use the bounds of the input after the transformation by $\mat{B}$ 
since the matrix may reverse the bounds. Eq.~\eqref{eq:of_bounds} can be solved 
using linear programming as it describes a polyhedral region in $\mathbb{R}^n$.
The second step is evaluating the progression of the system over time to calculate 
$\vec{x}_k$ to validate against this specification. For this step we require 
an initial set from which the system progresses. Obviously, this set $\vec{x}_0$ 
must satisfy the specification; otherwise, the system will be unsafe to begin with.
We accept a specification in the form $\mat{E}\mat{T}\vec{x}_0<\mat{f}$. 
The presence of $\mat{T}$ is because this will typically be given in the original 
state-space. Accelerating \eqref{eq:observer_LTI_cf}, we obtain:
%
\begin{align}
\label{eq:acc_observer_LTI_cf}
\vec{x}=&\mat{A}_t^k\vec{x}_0
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tp} \vec{r}_i
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tn}\left [\begin{array}{c}\vec{\nu}_2\\ \vec{\nu}_1\end{array}\right] : \mat{E}\mat{T} \vec{x}_0<\mat{f}
\end{align}
which would still require us to verify the system for every $k$ up to infinity. 
Instead we use abstract acceleration~\cite{cattaruzza2015unbounded}, which translates the formula into
%
\begin{align}
\label{eq:aa_observer_LTI_cf}
\hat{X}^\#
=&\mathcal{A}_t X_0+\mathcal{B}_{tp} R + \mathcal{B}_{tn} N\\
X_0 =&\left \{\vec{x} : \mat{E}\mat{T} \vec{x}_0<\mat{f} \right\}\nonumber\\
 R =&\{r : \mat{E}_r < \vec{f}_r \}\nonumber\\
 N=&\left \{ \left[\begin{array}{c}\vec{\nu}_2\\ \vec{\nu}_1\end{array}\right] : \nu_1 \in \left[-\frac{q1}{2}\ \ \frac{q1}{2}\right] \wedge \nu_2 \in \left[-\frac{q2}{2}\ \ \frac{q2}{2}\right]  \right \}\nonumber
\end{align}
%
where $\mathcal{A}_t=\bigcup_{k=1}^\infty \mat{A}^k, \mathcal{B}_{tp}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tp}, \mathcal{B}_{tn}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tn}$ are abstract matrices for the system in \eqref{eq:observer_LTI_cf}. It is worth noting that $r$ is a non-deterministic parametric input (\emph{ie} $\forall k>0, r_{k+1} =r_k$) whereas the set N is composed of non-deterministic variable inputs.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Controller synthesis}\label{sec:cof_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Now that we have explained the verification process, we may proceed to describe the synthesis procedure, which is illustrated in figure \ref{fig:CEGARIS}.
\begin{enumerate}
\item We start by preprocessing the dynamics. This can be computationally intensive and we don't want it to be part of the CEGIS loop.
\begin{enumerate}
\item Calculate $\mat{A}_o,\mat{B}_o,\mat{C}_o$ by rounding down the coefficients $*_{oij}=\mathcal{F}\langle I,F,E\rangle (*_{ij}) : *=\{a,b,c\}$
\item Calculate $\mat{T}$ and $\mat{W}_o^{-1}\mat{W}_{of}$ using equations \eqref{eq:rncf}\eqref{eq:rcf}\eqref{eq:to_cf}\eqref{eq:wnof}\eqref{eq:wof}
\item Extract the Plant's characteristic polynomial coefficients from the last row of $P_a=(\mat{T}\mat{A}\mat{T}^-1)_{n*}$.
\item Calculate the noise set $N$ from the quantizer resolutions and estimated round-off errors ($\frac{q_1}{2},\frac{q_2}{2},q_3$)
\item Produce an initial abstraction which corresponds to an unconstrained solution space.
\end{enumerate}
\item Next we synthesise a candidate controller and observer. Let $\mathcal{J}\langle I,F \rangle (P,\mat{T})$ be a program describing Jury's method.
\begin{enumerate}
\item Select a controller $\tilde{\mat{K}} \in \mathbb{R}\langle I,F\rangle^n$. This is achieved by solving a SAT formula $\mathcal{J}_K=\mathcal{J}\langle I,F,E \rangle (P_{a-k},\mat{T})$ that satisfies Jury's criteria for $P_{a-k}=z^n+\sum_{i=1}^n (a_i-k_i) z^{n-i} : k_i \in \mat{K} \wedge \mat{K}=\tilde{\mat{K}}\mat{T}$.
\item If there is no candidate solution we return UNSAT and exit the loop.
\end{enumerate}
\item Once we have a candidate solution, we perform a safety verification using abstract acceleration.
\begin{enumerate}
\item Calculate the abstract matrices $\mathcal{A}_t, \mathcal{B}_{tp},\mathcal{B}_{tn}$ for the candidate closed loop solution.
\item Evaluate $\hat{X}^\# \models \Psi$. If the verification holds we have a solution, and exit the loop.
\end{enumerate} 
\item If the candidate solution is not valid we refine the abstraction used by the synthesiser
\begin{enumerate}
\item Find the constraints that invalidate the property. These are the rows of $\mat{E} \in \Psi$ that intersect/segment $\hat{X}^\#$. Transforming these rows into the eigenspace of $\mat{A}_t$, and applying backward acceleration to obtain a set of counterexamples for the eigenvalues, which we define as $\Psi(\Lambda)$. This is a constraint in the spectrum (\emph{ie} frequency response transfer function) of the closed loop dynamics. 
\item Transform the soft constraints on the spectrum into soft constraints on the coefficients of the characteristic polynomial. 
$z^n+\sum_{i=1}^n[a_i-k_i]z^{n-i}=\prod_{i=1}^n (z-[\lambda_i]) : |[\lambda_i]|<1 \wedge [\lambda_i] \models \Psi_{\Lambda}$.
\item Pass the refined abstraction with the new contraints to the synthesiser, as well as a counterexample plant coefficients ($P_a$) and repeat the loop again. We note that the Abstraction refinement is a soft constraint on the proposed model which can increase or decrease the parameter space at any time, whereas the concrete counterexample passed to the synthesiser is monotonically decreasing. The first one improves speed of the tool whereas the second ensures convergence.
